Artin Group Presentations Arising from Cluster Algebras

### 2016-09-29 Paper.tex
Minor Changes
made simple edits from referee report

\title{Artin Group Presentations Arising from Cluster Algebras}
\author{Jacob Haley, David Hemminger, Aaron Landesman, Hailee Peck}
\email{jhaley3@nd.edu, david.hemminger@duke.edu, aaronlandesman@college.harvard.com, hpeck@millikin.edu}
\date{\today}
\keywords{Artin group, cluster algebra, diagram representations}

@@

The initial introduction of cluster algebras by Fomin and Zelevinsky was aimed at making further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite type} cluster algebras, those with a finite number of distinguished generators. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite type cluster algebras can be classified by Dynkin diagrams. \\

A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{initial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetric matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = \left\{ x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n \right\}$.
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = \left\{ x_1, \ldots, x_n \right\}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

@@

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra of finite type by the following set of rules:
\begin{prop}\cite{FZ03}
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows:
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident to $k$ (leaving the weights unchanged)
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and
$\pm \sqrt{c} \pm \sqrt{c'} = \sqrt{ab}$, 
where the sign before $\sqrt{c}$ (respectively, $\sqrt{c'}$) is positive if the arrows form an oriented cycle and negative otherwise. Then $\Gamma(B)$ changes as in Figure \ref{mutation}, taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{prop}

@@

\begin{defn}
Let $W_{\Gamma}$ be the group with generators $s_i, i = 1,2,\ldots, n$, subject to the following relations:
\begin{itemize}
\item{(R1)} $s_i^2 = e$ for all $i$
\item{(R2)} $\left(s_is_j\right)^{m_{ij}} = e$ for all $i \neq j$
\end{itemize}
Furthermore, for a chordless cycle $C : i_0 \rightarrow i_1 \rightarrow \cdots \rightarrow i_{d-1} \rightarrow i_0$ and for $a = 0,1,2,\ldots, d-1$, define \textbf{$r\left(i_a, i_{a+1}\right) = s_{i_a}s_{i_{a+1}} \cdots s_{i_{a+d-1}}s_{i_{a+d-2}} \cdots s_{i_{a+1}}$}.\\

\vspace{0.1cm}
Then we have the following relations:
\begin{itemize}
\item{(R3)(a)} If all the weights in the edges of $C$ are $1$, then $r(i_a, i_{a+1})^2 = e$
\item{(R3)(b)} If $C$ has some edges of weight $2$, then $r(i_a, i_{a+1})^k = e$ where $k = 4-w_a$ and $w_a$ is the weight of the edge $i_a\--\ i_{a-1}$
\end{itemize}
\end{defn}

@@

\begin{defn}
\label{chordlesscycle}
A {\it chordless cycle} of an unoriented graph $G$ is a connected subgraph $H \subset G$ such that the number of vertices in $H$ is equal to the number of edges in $H,$ and the edges in $H$ form a single cycle.
\end{defn}

\begin{prop}
\label{cycle_types}
\cite[Proposition 2.1]{BM13} Let $\Gamma$ be a diagram of finite type. Then, a chordless cycle in the unoriented graph of $\Gamma$ is cyclically oriented in $\Gamma$. Furthermore, the unoriented graph underlying the cycle must either be a cycle such that all edges have weight 1, a square with two opposite edges of weight 2 and two opposite edges of weight 1, or a triangle with two edges of weight 2 and one of weight 1, as pictured in Figure \ref{fig:cycle_types}.
\end{prop}

@@

In order to prove our main result, Theorem ~\ref{thm:main}, we must first define the Artin group associated to a finite type diagram. This definition will be similar to that made in \cite{BM13} at the beginning of Section 3, except that we shall not require relation $(R1),$ i.e., $s_i^2 = e.$ Since Artin Groups are very similar to Coxeter groups, with the caveat that the generators are not involutions, we will be able to use these modified relations to great effect.

@@

\begin{lem} \label{Sym-Lem}
Let $\Gamma$ be a diagram of finite type which contains a chordless cycle C:
\begin{equation*}
\begin{tikzcd}
i_0 \arrow{r} & i_1 \arrow{r} & \cdots \arrow{r} & i_{d-1} \arrow{r} & i_0
\end{tikzcd}
\end{equation*}
so that all edges have weight 1. Then if W is a group generated by $s_{1}, \dots, s_{n}$ satisfying the relations (T2) and $t(i_{a}, i_{a+1}) = e$ for some a $\in \{1, \dots, d\}$, all of the relations in (T3) hold for C.
\end{lem}

@@

\noindent is a group homomorphism as well.  By Lemma \ref{lem:op_homomorphism}, there exist two well-defined homomorphisms $\Delta\colon A_{\Gamma}\rightarrow A_\Gamma^{op}$ defined by $\Delta(s_i) = q_i^{-1}$ and $\Delta^\prime\colon A_{\left(\Gamma^\prime\right)^{op}} \rightarrow A_{\Gamma^\prime}$ defined by $\Delta^\prime(u_i) = r_i^{-1}$. We then have a homomorphism

$$\psi = \Delta^\prime\circ\varphi_{op}\circ\Delta\colon A(\Gamma)\rightarrow A(\Gamma^{op})\rightarrow A(\left(\Gamma^{op}\right)^{\prime}) \rightarrow A(\Gamma^{\prime})$$

@@

\item We have
\begin{align*}
t_3t_4^{-1}t_1t_4 &= s_3s_4^{-1}\underline{s_2s_1s_2^{-1}}s_4\\
&= s_3s_4^{-1}s_1^{-1}s_2s_1s_4\\
&= s_4^{-1}\underline{s_1^{-1}s_2s_1}s_4s_3 &\text{by} t(3,4)\\
&= s_4^{-1}s_2s_1s_2^{-1}s_4s_3\\
&= t_4^{-1}t_1t_4t_3\\
\end{align*}

>>>>

\title{Artin Group Presentations Arising from Cluster Algebras}
\author{Jacob Haley, David Hemminger, Aaron Landesman, Hailee Peck}
\email{jachaley@umich.edu, david.hemminger@duke.edu, aaronlandesman@college.harvard.com, hpeck@millikin.edu}
\date{\today}
\keywords{Artin group, cluster algebra, diagram representations}

@@

The initial introduction of cluster algebras by Fomin and Zelevinsky was aimed at making further strides in the areas of representation theory, Lie theory, and total positivity. Since then, the study of cluster algebras has provided a motivation for applications in various other areas of mathematics, including quiver representations. Of particular interest were \textit{finite type} cluster algebras, which we define precisely below. In the sequel to their introductory paper (\cite{FZ03}), Fomin and Zelevinsky introduce the concept of \textit{mutation equivalence} between diagrams, proving that a connected graph is mutation equivalent to an oriented Dynkin diagram if and only if all mutation equivalent graphs have edge weights not exceeding 3. In particular, this proves that finite type cluster algebras can be classified by Dynkin diagrams. \\

A \textit{cluster algebra} is an integral domain which can be generated by a set of elements called \textit{cluster variables} that satisfy certain exchange relations. Following the style of \cite{FZ02} and \cite{BM13}, we will define cluster algebras in terms of \textit{skew-symmetrisable} matrices (that is, a matrix $B$ such that there exists a diagonal matrix $D$ of the same size with $D_{ii} >0$ such that $DB$ is skew-symmetric). Let $\mathbb{F} = \mathbb{Q}(u_1, u_2, \ldots, u_n)$ be the field of rational functions in $n$ indeterminates over $\mathbb{Q}$. We will define an \textit{initial seed} for the cluster algebra to be a fixed pair $(\textbf{x}, B)$, where $\textbf{x} = \{x_1, \ldots, x_n\}$ is a free generating set of $\mathbb{F}$ and $B$ is an $n \times n$ skew-symmetrisable matrix. Define $x_k^{\prime} \in \mathbb{F}$ by the \textit{exchange relation}
\begin{displaymath}
x_k^{\prime}x_k = \prod_{B_{ik} > 0}{x_i^{B_{ik}}} + \prod_{B_{ik} < 0}{x_i^{-B_{ik}}}
\end{displaymath}
Then, given an initial seed $(\textbf{x}, B)$ and $k \in {1,2,\ldots,n}$, we can define a \textit{mutation} of the seed at $k$, denoted $\mu_k(\boldsymbol{x}, B) = (\textbf{x}^{\prime}, B^{\prime})$ where:
\begin{displaymath}
B_{ij}^{\prime} = \begin{cases} - B_{ij} & \mbox{ if } i = k \mbox{ or } j = k;\\
B_{ij} + \frac{|B_{ik}|B_{kj} + B_{ik}|B_{kj}|}{2} & \mbox{ otherwise. }\\
\end{cases}
\end{displaymath}
and $\textbf{x}^{\prime} = \left\{ x_1, x_2, \ldots, x_{k-1}, x_k^{\prime}, x_{k+1}, \ldots, x_n \right\}$.
Such a mutation or a sequence of such mutations generate \textit{seeds} which in turn generate all cluster variables in that, for each $\textbf{x} = \left\{ x_1, \ldots, x_n \right\}$ corresponding to a seed of the cluster algebra, the entries $x_i$ are the cluster variables. \\

@@

Just as we can define mutations of the seeds of a cluster variable, we can also define mutations of a diagram associated to a cluster algebra of finite type by the following set of rules:
\begin{prop}\cite{FZ03}
Let $B$ be a 2-finite skew-symmetrisable matrix. Then $\Gamma(\mu_k(B))$ is uniquely determined by $\Gamma(B)$ as follows:
\begin{itemize}
\item Reverse the orientations of all edges in $\Gamma(B)$ incident to $k$ (leaving the weights unchanged).
\item For any path in $\Gamma(B)$ of form $i \xrightarrow{a} k \xrightarrow{b} j$ (i.e. with $a,b$ positive), let $c$ be the weight on the edge $j \rightarrow i$, taken to be zero if there is no such arrow. Let $c'$ be determined by $c'\geq 0$ and
$\pm \sqrt{c} \pm \sqrt{c'} = \sqrt{ab}$, 
where the sign before $\sqrt{c}$ (respectively, $\sqrt{c'}$) is positive if the arrows form an oriented cycle and negative otherwise. Then $\Gamma(B)$ changes as in Figure \ref{mutation}, taking the case $c' = 0$ to mean no arrow between $i$ and $j$.
\end{itemize}
\end{prop}

@@

\begin{defn}
Let $W_{\Gamma}$ be the group with generators $s_i, i = 1,2,\ldots, n$, subject to the following relations:
\begin{itemize}
\item{(R1)} $s_i^2 = e$ for all $i$
\item{(R2)} $\left(s_is_j\right)^{m_{ij}} = e$ for all $i \neq j$
\end{itemize}
Furthermore, for a chordless cycle $C : i_0 \rightarrow i_1 \rightarrow \cdots \rightarrow i_{d-1} \rightarrow i_0$ and for $a = 0,1,2,\ldots, d-1$, define \textbf{$r\left(i_a, i_{a+1}\right) = s_{i_a}s_{i_{a+1}} \cdots s_{i_{a+d-1}}s_{i_{a+d-2}} \cdots s_{i_{a+1}}$}, where indices are considered $\bmod d.$\\

\vspace{0.1cm}
Then we have the following relations:
\begin{itemize}
\item{(R3)(a)} If all the weights in the edges of $C$ are $1$, then $r(i_a, i_{a+1})^2 = e$
\item{(R3)(b)} If $C$ has some edges of weight $2$, then $r(i_a, i_{a+1})^k = e$ where $k = 4-w_a$ and $w_a$ is the weight of the edge joining $i_a$ and $i_{a-1}$
\end{itemize}
\end{defn}

@@

\begin{defn}
\label{chordlesscycle}
A {\it chordless cycle} of an unoriented graph $G$ is a connected induced subgraph $H \subset G$ such that the number of vertices in $H$ is equal to the number of edges in $H,$ and the edges in $H$ form a single cycle.
\end{defn}

\begin{prop}
\label{cycle_types}
\cite[Proposition 9.7]{FZ03} Let $\Gamma$ be a diagram of finite type. Then, a chordless cycle in the unoriented graph of $\Gamma$ is cyclically oriented in $\Gamma$. Furthermore, the unoriented graph underlying the cycle must either be a cycle such that all edges have weight 1, a square with two opposite edges of weight 2 and two opposite edges of weight 1, or a triangle with two edges of weight 2 and one of weight 1, as pictured in Figure \ref{fig:cycle_types}.
\end{prop}

@@

In order to prove our main result, Theorem ~\ref{thm:main}, we must first define the Artin group associated to a finite type diagram. This definition will be similar to that made in \cite{BM13} at the beginning of Section 3, except that we shall not require relation $(R1),$ i.e., $s_i^2 = e.$ Since Artin groups are very similar to Coxeter groups, with the caveat that the generators are not involutions, we will be able to use these modified relations to great effect.

@@

\begin{lem} \label{Sym-Lem}
Let $\Gamma$ be a diagram of finite type which contains a chordless cycle C:
\begin{equation*}
\begin{tikzcd}
i_0 \arrow{r} & i_1 \arrow{r} & \cdots \arrow{r} & i_{d-1} \arrow{r} & i_0
\end{tikzcd}
\end{equation*}
so that all edges have weight 1. Then if A is a group generated by $s_{1}, \dots, s_{n}$ satisfying the relations (T2) and $t(i_{a}, i_{a+1}) = e$ for some a $\in \{1, \dots, d\}$, all of the relations in (T3) hold for C.
\end{lem}

@@

\noindent is a group homomorphism as well.  By Lemma \ref{lem:op_homomorphism}, there exist two well-defined homomorphisms $\Delta\colon A_{\Gamma}\rightarrow A_{\Gamma^{op}}$ defined by $\Delta(s_i) = q_i^{-1}$ and $\Delta^\prime\colon A_{\left(\Gamma^\prime\right)^{op}} \rightarrow A_{\Gamma^\prime}$ defined by $\Delta^\prime(u_i) = r_i^{-1}$. We then have a homomorphism

$$\psi = \Delta^\prime\circ\varphi_{op}\circ\Delta\colon A_\Gamma \rightarrow A_{\Gamma^{op}}\rightarrow A_{\left(\Gamma^{op}\right)^{\prime}} \rightarrow A_{\Gamma^{\prime}}$$

@@

\item We have
\begin{align*}
t_3t_4^{-1}t_1t_4 &= s_3s_4^{-1}\underline{s_2s_1s_2^{-1}}s_4\\
&= s_3s_4^{-1}s_1^{-1}s_2s_1s_4\\
&= s_4^{-1}\underline{s_1^{-1}s_2s_1}s_4s_3 &\text{by } t(3,4)\\
&= s_4^{-1}s_2s_1s_2^{-1}s_4s_3\\
&= t_4^{-1}t_1t_4t_3\\
\end{align*}
